---
title: "Random report"
author: 
  - name: Jane Doe
    orcid: 0000-0001-2345-6789
    corresponding: true
    email: Jane.Doe@example.org
    
---

<!-- 

# Instructions for the exercise

1. Move the file to the top level of your repository
2. Update the file _quarto.yml in your repository to include the report in your webpage
3. Load a dataset from https://github.com/Hezel2000/GeoCosmoChemDataAndTools/tree/main/csv
4. Update the YAML header of this file
5. Create a file `requirements.txt`on the top level of the repo
5. Update the CI/CD pipeline as instructed below
6. Create a short report, including plots created from the dataset (they can be interactive if you want -- the result is a webpage)
  a. Don't forget to update the `requirements.txt`with the libraries you call.  
  b. You are welcome to write and test any Python code elsewhere before including it in the report. 
  c. You might find the following pages useful for reference: 
      Figures: https://quarto.org/docs/authoring/figures.html
      Tables: https://quarto.org/docs/authoring/tables.html

The remainder of the document is a small example of how it might look like and provides some help to get started. Don't forget to compare the unrendered and the rendered version! 

By the way, to include Python code, write `python` instead of `r` in the curly brackets at the top of a code chunk. 

# Updating the CI/CD pipline

Add between lines 15 and 16: 

      - name: Install Python and Dependencies
        uses: actions/setup-python@v4
        with:
          python-version: '3.10'
          cache: 'pip'
      - run: pip install -r requirements.txt 

Make sure identations are correct! 

--> 

## The randomness of R

Because there must be some content for inspiration, let's see how random random numbers in R really are. 

## Uniformly distributed random numbers

Uniformly distributed random number means that all number have the same chance to get drawn from the pool. There are multiple methods to produce such numbers. 

The most common command to create a uniform distribution of random numbers in R is `runif()`, standing for **r**andom **unif**orm. There is also the command `sample()` which is a bit more complex and therefore not further discussed here. 

By default, `runif()` samples the range from 0 to 1: 

``` {r}

runif(1)

runif(10)

``` 

This range can be changed by providing additional arguments: 

``` {r}

runif(1, min = 6, max = 9)

runif(10, min = 4, max = 20)

``` 

Sometimes, integers are needed. This can be achieved with 

``` {r}

floor(runif(1, min = 1, max = 6))

``` 

With this information, let's see how random R throws a D20! 

``` {r}
#| layout-ncol: 2
#| fig-cap: 
#|   - "10 rolls"
#|   - "100 rolls"
#|   - "1000 rolls" 
#|   - "10000 rolls"
#| echo: false

hist(runif(10, min = 1, max = 20), breaks = seq(from = 0.5, to = 20.5), main = NULL)
hist(runif(100, min = 1, max = 20), breaks = seq(from = 0.5, to = 20.5), main = NULL)
hist(runif(1000, min = 1, max = 20), breaks = seq(from = 0.5, to = 20.5), main = NULL)
hist(runif(10000, min = 1, max = 20), breaks = seq(from = 0.5, to = 20.5), main = NULL)

```

And now let's check if the distribution is consistent within the same number of rolls: 

``` {r}
#| layout-ncol: 5
#| echo: false

for (i in seq_along(1:10)) {
  hist(runif(10, min = 1, max = 20), breaks = seq(from = 0.5, to = 20.5), main = NULL)
}

```

# Normally distributed random numbers

R cannot also create numbers from a normal distribution with the command `rnorm()` (= **r**andom **normal**). The syntax of the function is similar to `runif()` but instead of defining minimum and maximum of a function, we give the mean and the standard deviation ($sd$) of the normal distribution: 

``` {r}

rnorm(5)

rnorm(5, mean = 50, sd = 5)

```

Let's repeat the same test as above: 

``` {r}
#| layout-ncol: 2
#| fig-cap: 
#|   - "10 draws"
#|   - "100 draws"
#|   - "1000 draws" 
#|   - "10000 draws"
#| echo: false

hist(rnorm(10, mean = 50, sd = 5), main = NULL)
hist(rnorm(100, mean = 50, sd = 5), main = NULL)
hist(rnorm(1000, mean = 50, sd = 5), main = NULL)
hist(rnorm(10000, mean = 50, sd = 5), main = NULL)

```

And performance in repeated identical draws. However, this time we will draw 50 times because 10 is too few to make the normal distribution visible: 

``` {r}
#| layout-ncol: 5
#| echo: false

for (i in seq_along(1:10)) {
  hist(rnorm(50, mean = 50, sd = 5), main = NULL)
}

```


If you forgot how the standard deviation is calculated (and because I wanted to include a more complex formula for instructive purposes): 

$$
\sigma = \sqrt{\frac{\sum{(x_i-\mu)^2}}{N}} 
$$
with: 

| Symbol || Explanation |
|----:|----|:----|
|$\sigma$| = | Population standard deviation| 
|$N$| = | Size of population |
|x~i~| = | each value from the population |
|µ| = | mean of the population |

# Conclusions
As expected, the number from random distributions are pretty random 

Have fun while doing the exercise and don't hesitate to ask for help! 
